---
title: "Exploring Recovery and Catch-Sample Data"
author: "Eric C. Anderson"
date: "November 18, 2014"
output:
  html_document:
    fig_width: 10
    fig_height: 8
    toc: yes
---

```{r, echo=FALSE, include=FALSE}
library(knitr)
opts_knit$set(root.dir = '..')
```

```{r, echo=FALSE, include=FALSE}
# load libraries, etc
library(ggplot2)
library(lubridate)
library(dplyr)
library(reshape2)
source("R/rmis_cleaning.R")
```


## Initial Maneuvers

Here we look at some properties to make sure that we got everything we wanted:
```{r, cache=TRUE}
# load it in
chin_rec <- readRDS("data/chinook_recoveries.rds")
dim(chin_rec)
```
This accords well with what the RMIS data base tells us when we query All Recoveries for 2002-2012 on chinook:
```
Your All Recoveries query retrieved too many rows (1452480)
```

Let's grab the catch-sample data too:
```{r, cache=TRUE}
catch_samp <- readRDS("data/catch_sample.rds")
```

### Sizes of catch samples
First I want to look very crudely at the first-mark sample size vs the first-mark obs adclips:
```{r}
ggplot(data = catch_samp) + 
    geom_point(aes(x = mr_1st_sample_size, y = mr_1st_sample_obs_adclips, color = reporting_agency))

```

OK, that shows us that some of the samples are pretty much all ad-clipped, and some of them are far from being all
ad-clipped.

Another interesting question then is how many of the adclipped, sampled fish yielded recovered and decoded coded
wire tags, broken out by Electronic versus Visual detection:
```{r}
ggplot(data = catch_samp) + 
    geom_abline(intercept = 0, slope = 1) +
    geom_point(aes(x = mr_1st_sample_obs_adclips, y = number_recovered_decoded, color = reporting_agency)) + 
    facet_wrap(~ detection_method)
```

OK, since there are so many reporting agencies, and the colors are hard to discern, it would be nice to know how many 
of the points come from each of the agencies.  Since many rows have been dropped due to missing data, I would
like to reflect that.
```{r}
ff <- filter(catch_samp, !is.na(mr_1st_sample_obs_adclips), !is.na(number_recovered_decoded))
ggplot(ff, aes(x = reporting_agency, fill = reporting_agency)) + geom_bar() + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

OK, that is interesting.  It suggests to me that CDFW and CDFO aren't really listing the number of observed adclips 
or the number recovered decoded.  So, what if we don't filter and just look at who has the most records.
```{r}
ggplot(catch_samp, aes(x = reporting_agency, fill = reporting_agency)) + geom_bar() + 
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

All right, that just shows us that CDFO and CDFW just don't have that many catch-sample records.  That might be 
because each of their catch samples includes more fish.

So, what we will want to do here is actually count up the number of fish from each agency in the recovery data
base (splitting them out between status 1 recoveries and otherwise)
```{r}
chin_rec$tag_status <- factor(tag_status_levels()[as.character(chin_rec$tag_status)], levels = unname(tag_status_levels()))
tag_nums <- chin_rec %>%
  group_by(run_year, reporting_agency, tag_status, detection_method) %>% 
  tally()
ggplot(tag_nums, aes(x = reporting_agency, y = n, fill = tag_status)) +
  geom_bar(stat = "identity") +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) + 
  facet_wrap(~ run_year + detection_method, ncol = 2)
```

That is a nice graphic.  It shows that ADFG and CDFO have real problems with ad-clipped fish that don't have a CWT,
but only in fisheries that don't use electronic detection (E = electronic, V = visual). OK.  That makes sense.

So, I still feel like I need to try to figure out how the numbers in the catch_sample file relate to the
tallies that I could make in the recovery data base.


### What is a catch-sample anyway?

Are there multiple catch-samples with the same code (just on different dates, etc?)
```{r}
catch_samp %>% 
  group_by(catch_sample_id) %>%
  tally() %>%
  select(n) %>%
  table()
```