---
title: "Summarize for MCMC"
author: "Eric C. Anderson"
date: "Jan 4, 2014"
output:
  html_document:
    fig_width: 10
    fig_height: 8
    toc: yes
---

```{r, echo=FALSE, include=FALSE}
library(knitr)
opts_knit$set(root.dir = '..')
```

```{r, echo=FALSE, include=FALSE}
# load libraries, etc
library(ggplot2)
library(lubridate)
library(dplyr)
library(reshape2)
library(stringr)
source("R/rmis_cleaning.R")
```


## Objective

The objective here is make a nice summary for each `catch_sample_id` which enumerates the numbers of different categories
described in `tex/sketches/cwt-ppn-estimation.tex`.

To get that, we are going to need to draw from the recovery data base and also from the catch_sample data base.  

Let's start with getting what we can from the recovery data base.

### Recovery Data Base Operations

First, make some new fields that relate to the variables we want:
```{r}
rec <- tbl_df(readRDS("data/chinook_recoveries.rds"))
rec <- rec %>% filter(run_year == 2012) # filter it down to 2012

# one of the first things we will do here is make a catch_sample_id that should
# be unique.  The data base is pretty screwed up that way.
rec <- rec %>% mutate(unique_catch_sample_id = paste(catch_sample_id, reporting_agency, run_year, sep = "--"))

# add leading zeroes onto the mark codes
rec$recorded_mark <- str_pad(rec$recorded_mark, 4, "left", "0")

# then get ready to switch the marks to our variables
mycodes <- vector()
mycodes["9"] <- "unknown"
mycodes["5"] <- "yes"
mycodes["0"] <- "no"

# and do it:
rec <- rec %>% mutate(ad_clipped = mycodes[str_sub(recorded_mark, 1, 1)])


# now, assume all electronically detected fish were beep fish (must be the case, I believe)
rec <- rec %>% mutate(beep = ifelse(detection_method == "E", "yes", "irrelevant"))

# now, get the tag statuses:
tagstats <- vector()
tagstats[1] <- "cwt"
tagstats[2] <- "no_tag"
tagstats[c(3,4,7)] <- "no_read" 
tagstats[8] <- "unknown"  # if head is not processed then it gets unknown
tagstats[9] <- "awt"

# and we make our cwt_status column here
rec <- rec %>% mutate(cwt_status = tagstats[tag_status])
```

OK, with that done, we have `detection_method` `beep`, `ad_clipped`, and `cwt_status` as desired.  Let's count up all the
different levels of those:
```{r}
rec %>% group_by(detection_method, beep, ad_clipped, cwt_status) %>% tally()
```

So, that is a summary of the observed data in the recovery data base for 2012.  Now, we need to figure out how to combine
the catch_sample data with all of that.  That we do in the next section.

### Adding in the catch-sample data

First thing, we get the catch-sample data and whittle it down to only those samples that led to 2012 recoveries.

```{r}
cs <- tbl_df(readRDS("data/catch_sample.rds")) %>%
  mutate(unique_catch_sample_id = paste(catch_sample_id, reporting_agency, catch_year, sep = "--"))

cs <- cs %>% filter(unique_catch_sample_id %in% rec$unique_catch_sample_id)
```

Later on we will be filtering out to only marine fishery `unique_catch_sample_id`'s which is good to keep in mind.  Since this stuff
now probably includes a lot of other stuff.  

#### Visually sampled fisheries
For visually sampled fisheries, I think it will end up being prett easy.  All we need to do is grab the total number of
non-adclipped fish, and we assign cwt_status of unknown to all of them.  The necessary information is held in the
`mr_` fields.  
```{r}
vsamp <- cs %>% filter(detection_method == "V") %>% 
  mutate(n = mr_1st_partition_size - mr_1st_sample_obs_adclips,
         cwt_status = "unknown",
         ad_clipped = "no",
         beep = "irrelevant") %>%
  select(unique_catch_sample_id, adclip_selective_fishery, number_sampled, mr_1st_partition_size, detection_method, beep, ad_clipped, cwt_status, n)
  
```
That is ready to be rbinded right onto the recovery data summaries, except that I included a few extra columns there because
I was curious about how there were no ad-clipped fish at all.  Perhaps those are freshwater or at-hatchery recoveries that will
be dropped anyway.

#### Electronically sampled fisheries

These are somewhat more complicated because of the several differing ways that they do mark sampling. The sampling for marks
is either dependent or independent of the beep-status of the fish.  If sampling is independent, then the `mr2_` fields are
NAs.  

Let's first deal with the dependent case.  Here are the no-beep fish.
```{r}
esamp_dep_no_beep <- cs %>%
  filter(detection_method == "E", !is.na(mr_2nd_partition_size) & !is.na(mr_2nd_sample_known_ad_status)) %>% # note that testing mr_2nd_sample_known_ad_status here is necessary because ODFW screwed up and put zeroes in instead of leaving  mr_2nd_partition_size absent, and the maintainers of the data base didn't catch it.  So lame, and such a waste of my time trying to figure out what is going on...GRRRRR...
  mutate(yes = mr_2nd_sample_obs_adclips, 
         no = mr_2nd_sample_known_ad_status - mr_2nd_sample_obs_adclips,
         unknown = mr_2nd_partition_size - mr_2nd_sample_known_ad_status,
         beep = "no", 
         cwt_status = "unknown") %>%
  select(unique_catch_sample_id, detection_method, beep, cwt_status, yes, no, unknown) %>%
  melt(measure.vars = c("yes", "no", "unknown"), variable.name = "ad_clipped", value.name = "n") %>%
  tbl_df %>%
  arrange(unique_catch_sample_id, ad_clipped) %>%
  select(unique_catch_sample_id:beep, ad_clipped, cwt_status, n)

esamp_dep_no_beep
```
and here are the beep-fish:
```{r}
esamp_dep_beep <- cs %>%
  filter(detection_method == "E", !is.na(mr_2nd_partition_size) & !is.na(mr_2nd_sample_known_ad_status)) %>%
  mutate(yes = mr_1st_sample_obs_adclips, 
         no = mr_1st_sample_known_ad_status - mr_1st_sample_obs_adclips,
         unknown = mr_1st_partition_size - mr_1st_sample_known_ad_status,
         beep = "yes", 
         cwt_status = "unknown") %>%
  select(unique_catch_sample_id, detection_method, beep, cwt_status, yes, no, unknown) %>%
  melt(measure.vars = c("yes", "no", "unknown"), variable.name = "ad_clipped", value.name = "n") %>%
  tbl_df %>%
  arrange(unique_catch_sample_id, ad_clipped) %>%
  select(unique_catch_sample_id:beep, ad_clipped, cwt_status, n)

esamp_dep_beep
```
These latter ones are a little awkward because I suspect many of these are aleady going to be reported in the recovery file. I still need to figure out what to do about that.  

Finally, we have the (few) situations where sampling for marks occurs independently of the electronic sampling.  It is not clear
whether some of the fish sampled might be the same as the ones that had CWTs recovered from or not.  It's unclear the way
they have set it up.  But, I am just going to go ahead and assume that the fish investigated for ad-clips here are
entirely distinct from those that were electronically sampled.  This might not be correct, but there is no other way to
deal with it.

Let's pick them out:
```{r}
esamp_indep <- cs %>%
  filter(detection_method == "E", is.na(mr_2nd_partition_size) | is.na(mr_2nd_sample_known_ad_status)) %>%
  mutate(yes = mr_1st_sample_obs_adclips, 
         no = mr_1st_sample_known_ad_status - mr_1st_sample_obs_adclips,
         unknown = mr_1st_partition_size - mr_1st_sample_known_ad_status,
         beep = "irrelevant", 
         cwt_status = "unknown") %>%
  select(unique_catch_sample_id, detection_method, beep, cwt_status, yes, no, unknown) %>%
  melt(measure.vars = c("yes", "no", "unknown"), variable.name = "ad_clipped", value.name = "n") %>%
  tbl_df %>%
  arrange(unique_catch_sample_id, ad_clipped) %>%
  select(unique_catch_sample_id:beep, ad_clipped, cwt_status, n)

esamp_indep
```


### catch-samp vs recovery for the electronically sampled beep fish

Here I take a quick look at the correspondence (if any) between `esamp_dep_beep` and what
we find in the recovery sample.
```{r}
# first, add up all the ad-clipped or not fish from the beep samples in the recovery file
tmp <- rec %>% 
  filter(detection_method == "E") %>%
  group_by(unique_catch_sample_id, detection_method, beep, ad_clipped) %>%
  tally() %>%
  filter(beep == "yes") %>%
  rename(nrec = n)

# then inner join that to the esamp_dep_beeps:
joined <- inner_join(esamp_dep_beep %>% select(-cwt_status), tmp) %>% 
  arrange(unique_catch_sample_id, detection_method, beep, ad_clipped)

dim(joined)

# now, see which ones don't match up
mism <- joined %>%
  filter(n != nrec)
```

Here is interesting thing to notice.  When it comes to ad-clipped fish 