---
title: "Variability in PBT tag rates"
author: "Eric C. Anderson"
date: "Feb 6, 2015"
output:
  html_document:
    fig_width: 10
    fig_height: 8
    toc: yes
---

```{r, echo=FALSE, include=FALSE}
library(knitr)
opts_knit$set(root.dir = '..')
```

```{r, echo=FALSE, include=FALSE}
# load libraries, etc
library(ggplot2)
library(lubridate)
library(dplyr)
library(reshape2)
library(stringr)

```


## Introduction

For PBT to function in the way that the CWT system
does requires that the PBT tagging rate
be known (or at least estimated).  One issue that arises in this context is 
the extra variability in tagging rate that might result from differential
survival of different families in the hatchery.  The variance in the PBT
tagging rate depends on the variance in family size in the hathery, as well
as on the pattern of mating adopted by the hatchery, and the fraction of parents
that were successfully genotyped. 

In order to predict the variance in
PBT tagging rate that might be encountered, this document presents the 
rationale and code behind a system of simulations that are paramterized
using the data that we currently have available at different hatcheries.
The inputs to this procedure are:

1. A data frame holding the recovered parentage assignments at the hatchery
over the course of PBT operations at it.
2. A vector of successful genotyping rates.

From these inputs, the method estimates the parameter of a Compound Dirichlet
Multinomial Distribution that best fits the pedigree data, and then it uses
these parameters to simulate fractions of offspring in a very large pool
of offspring (equivalent to a release group) and then, after assigning
the parents of these offspring to tagged or untagged groups, measures the 
tagging rate.  It does this repeatedly to assess the variance in tagging 
rate.


## Pedigree data

An example of the pedigree data appears below.
```{r read-data}
x <- tbl_df(read.table("data/FRH_trios.txt")) %>%
  mutate(brood_year = str_match(pa, "_(2[0-9]{3})$")[,2])

x
```

Now, for the purposes of estimating the parameters of the CDM distribution, we can condense this
down to the counts of full-sib families of different sizes (Note, we assume that we can capture
most of what we need to without a model that posits individual parent contributions to fitness
in half-sib families...)

```{r condense}
num_offs <- x %>% 
  group_by(brood_year, pa, ma) %>%
  tally() %>% rename(num_offspring = n) %>%
  ungroup

fam_sizes <- num_offs %>%
  group_by(brood_year, num_offspring) %>%
  tally()

head(fam_sizes, n = 15)
  
```

The first thing we see from this is that we have 0-censored data---because we don't have individual
mating records in this hatchery, we don't know which individuals were mated together, so we don't have
an exact count of how many parent pairs there were that didn't produce any offspring. However, knowing the
fraction of sampled parent pairs, we can at least come up with an estimate of how many matings there
may have been that we wouldn't be able to recover through parentage, anyway. 



## Estimation of alpha and K

Since we are interested in the variance in family size, we will use that as a criterion for fitting
the model. We will also fit the number parent pairs with at least one offspring.  We will do this
by simulation.  First we need something that simulates the CDM and computes the variance of the 
non-zero family sizes from it.  

First let us see what the mean number and sd of (non_zero) family size is each year
and also the number of families with at least one offspring detected.

```{r}
means_etc <- num_offs %>% 
  group_by(brood_year) %>%
  summarise(mean = mean(num_offspring), sd = sd(num_offspring), size = sum(num_offspring), non_zero = n())

means_etc
```

We can think of those values as living in the cartesian plane:
```{r}
ggplot(means_etc, aes(x = sd, y = non_zero, colour = brood_year)) + geom_point(size = 5)
```

And, for each of them, we can simulate a CDM with n equal to non-zero, and trying
different values of a and K till we get something that works pretty well.  
```{r}
source("R/pbt-tagging-fract-funcs.R")

# then fiddle around with some things like this:
cdm_stats(simCDM(1000, size = 1024, a = 1.0, K = 600))

```
And what we find is that it is very hard to get the sorts of numbers that we see in the
observed data without cranking the number of parent pairs with no offspring up
higher than is reasonable.

It might be better to do more generic simulations just assuming the CDM and some
simple idea like Ne/N ratio.  


## Doing it via a gamma and urn model

If we are interested in the Ne/N ration and call that L.  Then, in a constant sized
population we can get that L by choosing the scale of a gamma shape parameter
to be $(L-1/N)/(1-L)$.

So, if we are talking about 50 families, and we want to think about them creating
a random number of offspring each with an average of 2, as if they were a perfect Poisson, then we would do:

```{r}
fsizes <- rpois(50, 2)
```

We could envisage that as a two-stage sampling process where we first simulate a
bunch of juveniles, and then randomly sample from them.  

The poisson variance in family size would be like everyone making an equal and very
number of juveniles and then randomly sampling from them.  But we could lump that extra
variance needed to reduced the effective size into a negative binomial distribution on junveniles.

That sort of hierarchy would look like this
```{r}
K <- 250 # number of mated pairs
f <- 4000 # average fecundity of  female chinoook
L <- .3 # ratio of Ne to N
shape = (L - 1 / K) / (1 - L)
scale = f / shape

g <- rgamma(K, shape = shape, scale = scale)
poi <- rpois(length(g), g) # neg bin distributed juveniles
adu <- rmultinom(1, size = K, prob = poi/sum(poi))
wf <- rmultinom(1, size = K, prob = rep(1/K, K))
```

But I think that is totally crackpot.  It will probably be easier and better to
think of the parent generation of size K giving rise to a juvenile offspring 
population of size K*f, and then twiddle the $\lambda = N_e/N$ knob between
different reasonable values.  In that case,  let's make a function:
```{r}
hsim <- function(K, f, L) {
  s = (K * (1 - L)) / (L * K * f - 1) # stochastic replacement
  shape = 1 / s
  scale = f / shape
  
  g <- rgamma(K, shape = shape, scale = scale)
  poi <- rpois(length(g), g) # neg bin distributed juveniles
  poi
  
}

hsim(25, 4000, .9)
hsim(25, 4000, .01)
```

That does not look right to me either.  That is far too little variation.  It 
might be better to just parameterize it in terms of the variance or even just
show a histogram for the distribution of juvenile offspring number.


So let us do that:
```{r}
K <- 25

b <- rnbinom(n = K, mu = 4000, size = 10)
```